---
title: "Modeling"
author: "Ayush Desai"
date: "`r Sys.Date()`"
output: html_document
---

# Load Packages

```{r setup, include=FALSE}
# sets global chunk options: show code, hide messages and warnings
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(rstanarm)   # bayesian regression models via Stan
library(tidyverse)  # data wrangling and plotting
library(bayesplot)  # posterior visualisation helpers
library(loo)        # leave-one-out cross-validation for model comparison

# uses all available CPU cores for parallel MCMC sampling
options(mc.cores = parallel::detectCores())
```

# Load Data

```{r}
df <- read_csv("airbnb_clean.csv") %>%
  mutate(
    # strips dollar signs and commas, then casts price to numeric
    price_num    = as.numeric(str_replace_all(price, "[$,]", "")),
    # log-transforms price to reduce right skew before modeling
    log_price    = log(price_num),
    # converts borough and room type strings to factors for modeling
    borough      = factor(neighbourhood_group_cleansed),
    room_type    = factor(room_type),

    # computes median log-price within each neighbourhood
    # captures micro-location value better than borough alone
    neigh_median_price = ave(
      log(as.numeric(str_replace_all(price, "[$,]", ""))),
      neighbourhood_cleansed,
      FUN = function(x) median(x, na.rm = TRUE)
    ),

    # measures how far a listing's price sits above or below its neighbourhood median
    price_vs_neigh = log_price - neigh_median_price,

    # flags listings that show high availability but almost no review activity
    # also catches zero-review listings from hosts with many properties
    ghost = case_when(
      availability_365 > 300 & (is.na(reviews_per_month) | reviews_per_month < 0.1) ~ 1L,
      reviews_per_month == 0 & calculated_host_listings_count > 3 ~ 1L,
      TRUE ~ 0L
    ),

    # log1p-transforms skewed count variables to reduce outlier leverage
    log_reviews      = log1p(number_of_reviews),
    log_host_count   = log1p(calculated_host_listings_count),

    # averages the five review sub-scores into a single completeness signal
    review_complete = rowMeans(
      select(., review_scores_rating, review_scores_cleanliness,
             review_scores_communication, review_scores_location,
             review_scores_value),
      na.rm = TRUE
    )
  ) %>%
  # removes rows with missing price, extreme prices, or missing key fields
  filter(
    !is.na(log_price),
    price_num > 10, price_num < 2000,
    !is.na(accommodates),
    !is.na(number_of_reviews)
  )

# prints dataset size and how many listings were ghost-flagged
cat("Listings:", nrow(df), "| Ghost-flagged:", sum(df$ghost), "\n")
```

# Price Prediction

## Model

Key improvements over v1:
- `neigh_median_price` replaces raw borough — neighbourhood-level location signal is much tighter
- `log_reviews` instead of raw count — reduces leverage from high-review outliers
- `review_complete` combines all five review subscores into one signal
- `price_vs_neigh` dropped from price model (would create circularity — it's derived from price)

```{r}
# fits a bayesian linear regression predicting log(price)
# normal(0, 0.5) shrinkage prior keeps coefficients conservative
# exponential(1) prior on sigma nudges residual variance toward smaller values
price_model <- stan_glm(
  log_price ~ neigh_median_price + room_type + accommodates +
              log_reviews + review_complete,
  data            = df,
  family          = gaussian(),
  prior_intercept = normal(5.0, 1.0),  # centred near e^5 ≈ $148, a reasonable average nightly price
  prior           = normal(0, 0.5),
  prior_aux       = exponential(1),
  chains = 4, iter = 2000, warmup = 500,
  seed = 42
)
```

## Posterior Predictive Check

```{r}
# overlays 50 simulated price distributions on the observed one to assess model fit
pp_check(price_model, nreps = 50) +
  labs(title = "Price Model — Posterior Predictive Check", x = "log(price)") +
  theme_minimal()
```

## Coefficient Credible Intervals

```{r}
# extracts 95% credible intervals for each coefficient and plots them as segments
# a segment that doesn't cross zero indicates a credibly non-zero effect
posterior_interval(price_model, prob = 0.95) %>%
  as.data.frame() %>%
  rownames_to_column("param") %>%
  filter(param != "(Intercept)") %>%
  ggplot(aes(y = reorder(param, `97.5%`))) +
  geom_segment(aes(x = `2.5%`, xend = `97.5%`, yend = param),
               linewidth = 1.2, color = "darkblue") +
  # dashed line at zero marks the boundary between positive and negative effects
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "95% Credible Intervals — Price Predictors",
       x = "Effect on log(price)", y = NULL) +
  theme_minimal()
```

## Predict Price for a New Listing

```{r}
# defines a hypothetical listing: manhattan median neighbourhood, entire home, 2 guests, 30 reviews, 4.8 stars
new_listing <- data.frame(
  neigh_median_price = median(df$neigh_median_price[df$borough == "Manhattan"], na.rm = TRUE),
  room_type          = factor("Entire home/apt", levels = levels(df$room_type)),
  accommodates       = 2,
  log_reviews        = log1p(30),
  review_complete    = 4.8
)

# draws from the posterior predictive distribution and back-transforms from log scale
pred_draws <- exp(posterior_predict(price_model, newdata = new_listing)[, 1])

# prints the median predicted price and the central 90% credible interval
cat("Median predicted price: $", round(median(pred_draws), 2), "\n")
cat("90% Credible Interval:  $", round(quantile(pred_draws, 0.05), 2),
    "– $", round(quantile(pred_draws, 0.95), 2), "\n")

# plots the full distribution of predicted prices with the median marked in red
tibble(price = pred_draws) %>%
  ggplot(aes(x = price)) +
  geom_histogram(bins = 60, fill = "darkblue", alpha = 0.8) +
  geom_vline(xintercept = median(pred_draws), color = "black", linewidth = 1) +
  scale_x_continuous(labels = scales::dollar, limits = c(0, 800)) +
  labs(title = "Posterior Predictive Distribution — Nightly Price",
       subtitle = "Manhattan · Entire home · 2 guests · 4.8★",
       x = "Predicted Price", y = "Draw count") +
  theme_minimal()
```

## Model Fit (LOO-CV)

```{r}
# runs leave-one-out cross-validation to estimate out-of-sample predictive accuracy
loo(price_model)
```

# Ghost Listing Detection

## Model

Key improvements over v1:
- `log_host_count` replaces raw count — reduces influence of hosts with 50+ listings
- `price_vs_neigh` added — ghost listings are often priced way above their neighbourhood (listed but not meant to book)
- `review_complete` added — incomplete review profiles correlate with inactive listings
- `log_reviews` added — a listing with zero reviews after months of availability is a strong ghost signal

```{r}
# selects only the columns needed for the ghost model and drops rows with any missing values
ghost_vars <- c("ghost", "borough", "room_type", "accommodates",
                "log_host_count", "price_vs_neigh", "review_complete", "log_reviews")

df_ghost <- df %>% drop_na(all_of(ghost_vars))
cat("Ghost model rows:", nrow(df_ghost), "\n")

# fits a bayesian logistic regression predicting the probability a listing is a ghost
# prior_intercept of -3 reflects that most listings are not ghosts (~5% base rate)
ghost_model <- stan_glm(
  ghost ~ borough + room_type + accommodates +
          log_host_count + price_vs_neigh +
          review_complete + log_reviews,
  data            = df_ghost,
  family          = binomial(link = "logit"),
  prior_intercept = normal(-3, 1),
  prior           = normal(0, 0.5),
  chains = 4, iter = 2000, warmup = 500,
  seed = 42
)

# computes each listing's posterior mean ghost probability and attaches it to the data
df_ghost$ghost_prob <- colMeans(posterior_epred(ghost_model))
```

## Ghost Rates by Borough

```{r}
# summarises ghost flag counts and mean posterior probabilities by borough
ghost_summary <- df_ghost %>%
  group_by(borough) %>%
  summarise(
    n           = n(),
    n_flagged   = sum(ghost),
    pct_flagged = mean(ghost) * 100,
    avg_prob    = mean(ghost_prob),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_prob))

print(ghost_summary)

# plots mean ghost probability per borough with raw flagged counts as labels
ggplot(ghost_summary, aes(x = reorder(borough, avg_prob), y = avg_prob)) +
  geom_col(fill = "tomato", alpha = 0.85) +
  geom_text(aes(label = paste0(n_flagged, " flagged")), hjust = -0.1, size = 3.5) +
  coord_flip() +
  scale_y_continuous(limits = c(0, 0.5), labels = scales::percent) +
  labs(title = "Posterior Ghost Probability by Borough",
       x = NULL, y = "Mean P(ghost)") +
  theme_minimal()
```

## Ghost Coefficient Credible Intervals

```{r}
# plots 95% credible intervals for each ghost model coefficient on the log-odds scale
# a segment that doesn't cross zero indicates a credibly non-zero effect on ghost probability
posterior_interval(ghost_model, prob = 0.95) %>%
  as.data.frame() %>%
  rownames_to_column("param") %>%
  filter(param != "(Intercept)") %>%
  ggplot(aes(y = reorder(param, `97.5%`))) +
  geom_segment(aes(x = `2.5%`, xend = `97.5%`, yend = param),
               linewidth = 1.2, color = "black") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkblue") +
  labs(title = "95% Credible Intervals — Ghost Listing Predictors",
       x = "Effect on log-odds of being a ghost", y = NULL) +
  theme_minimal()
```

## Export Flagged Listings

```{r}
# filters to listings with posterior ghost probability above 15% and selects display columns
ghost_flagged <- df_ghost %>%
  filter(ghost_prob > 0.15) %>%
  select(id, name, neighbourhood_group_cleansed, neighbourhood_cleansed,
         room_type, price_num, availability_365,
         reviews_per_month, number_of_reviews, ghost_prob) %>%
  arrange(desc(ghost_prob))

# writes flagged listings to csv and prints the export count
write_csv(ghost_flagged, "ghost_listings_flagged.csv")
cat("Exported", nrow(ghost_flagged), "high-probability ghost listings.\n")
```